A Convolutional Neural Network for Outsole Recognition\tnoteref{t1}

### 2019-06-12 paper/paper.Rnw
Add brand ideas

\mt{The model base of VGG16 is made up of five convolutional blocks alternated with max pooling layers. Each convolutional block is made up of two or three convolutional layers, and each layer is made up of hundreds of filters, such that each of the five convolutional blocks has between 128 and 1,536 filters. Strictly speaking, a filter is a set of numerical weights that, when applied to an image, quantifies the presence of the specific shapes and/or colors that the filter is trained to detect. \fix{Necessary? (Or just skip...): Visualizing a filter then refers to visualizing the features to which the filter provides its maximum response.} \autoref{fig:vgg16layers} shows a selection of filters from VGG16, with each row corresponding to one of the five convolutional blocks. Filters in earlier convolutional blocks detect simple features, such as colors and blobs, while later filters detect more complex combinations of features. When considered together, the series of filters perform the process of feature detection, such that a single image is made up of some combination of the thousands of features that are detected by filters. Although the filters of VGG16 have been trained and optimized on the ImageNet Data, the large number of filters span a wide variety of shapes and features} \fix{which are still useful/relevant to classifying outsoles. In a sense, this is much like teaching humans to classify objects they have never seen before: an object may be unfamiliar, but our eyes can still detect its color, shape, and texture because we have learned these ideas from other sources.}

@@

Although transfer learning substantially reduces the amount of data required to train a CNN by several orders of magnitude, labeled images are still \fix{difficult or time-consuming} to generate relative to the amount of data necessary to train even a small neural network adequately. Thus, to make the most efficient use of the existing labeled data, we enlarge the training data using a process called image augmentation \citep{krizhevskyImageNetClassificationDeep2012}. Augmentation is the transformation of original input data using image operations such as cropping, zoom, skew, rotation, and color balance modification in order to distort or alter the image while maintaining the essential features corresponding to the label. This process reduces the potential for overfitting the model to the specific set of image data used during the training process, and also increases the amount of data available for training. Examples of pre- and post-augmentation images are shown in \autoref{fig:augmented}.

@@

Overfitting is when a model learns the training data so well that its understanding of the categories becomes specific to the training cases, which, in turn, leads to poor prediction of new images that were not in the training set. In training, overfitting is indicated when the validation loss begins to increase after reaching a global minimum. Alternately, underfitting occurs if the validation accuracy is still increasing when model optimization is terminated (for instance, between epochs 1 and 10 in \autoref{fig:training-accuracy}), because model performance is still improving with continued training. In \autoref{fig:training-accuracy}, validation accuracy levels off after epoch 20, and validation loss has not yet begun to increase at epoch 30, indicating that the model optimization process was halted at an appropriate epoch.

@@

%\autoref{fig:confusion_matrix} shows systematic errors in class predictions, but does not necessarily suggest \emph{why} the confusion is taking place. To assess the reasons for model failure, we look at the images to \fix{see what the model is seeing.}

@@

@@

>>>>

\mt{The model base of VGG16 is made up of five convolutional blocks alternated with max pooling layers. Each convolutional block is made up of two or three convolutional layers, and each layer is made up of hundreds of filters, such that each of the five convolutional blocks has between 128 and 1,536 filters. Strictly speaking, a filter is a set of numerical weights that, when applied to an image, quantifies the presence of the specific shapes and/or colors that the filter is trained to detect. \fix{Necessary? (Or just skip...): Visualizing a filter then refers to visualizing the features to which the filter provides its maximum response.} \autoref{fig:vgg16layers} shows a selection of filters from VGG16, with each row corresponding to one of the five convolutional blocks. Filters in earlier convolutional blocks detect simple features, such as colors and blobs, while later filters detect more complex combinations of features. When considered together, the series of filters perform the process of feature detection, such that a single image is made up of some combination of the thousands of features that are detected by filters. Although the filters of VGG16 have been trained and optimized on the ImageNet Data, the large number of filters span a wide variety of shapes and features} \fix{which are still useful/relevant to classifying outsoles. In a sense, this is much like teaching humans to classify objects they have never seen before: an object may be unfamiliar, but our eyes can still detect its color, shape, and texture because we have learned these ideas from other sources.} \fix{Say somewhere: Max pooling layers aggregate the features in one block before passing them to the next block, which helps reduce the computational load of training the model and helps to generalize the relationship between features.}

@@

Transfer learning substantially reduces the amount of data required to train a CNN by several orders of magnitude, but labeled images are still difficult to generate on a large scale. Thus, to make the most efficient use of the existing labeled data, we enlarge the training data using a process called image augmentation \citep{krizhevskyImageNetClassificationDeep2012}. Augmentation is the transformation of original input data using image operations such as cropping, zoom, skew, rotation, and color balance modification in order to distort or alter the image while maintaining the essential features corresponding to the label. This process reduces the potential for overfitting the model to the specific set of image data used during the training process, and also increases the amount of data available for training. Examples of pre- and post-augmentation images are shown in \autoref{fig:augmented}.

@@

Overfitting is when a model learns the training data so well that its understanding of the categories becomes specific to the training cases, which, in turn, leads to poor prediction of new images that were not in the training set. During the training process, overfitting is indicated when the validation loss begins to increase after reaching a global minimum. Alternately, underfitting occurs if the validation accuracy is still increasing when model optimization is terminated (for instance, between epochs 1 and 10 in \autoref{fig:training-accuracy}), because model performance is still improving with continued training. In \autoref{fig:training-accuracy}, validation accuracy levels off after epoch 20, and validation loss has not yet begun to increase at epoch 30, indicating that the model optimization process was halted at an appropriate epoch.

@@

@@

One possible explanation for the poor performance of some categories is the amount of variability within the shapes. \fix{For example, as seen in fig, a large proportion of bowties come from Birks; thus, the model learns bowties from Birks well, which leads to more accurate predictions of new bowtie images that come from Birks. There is no similar dominance in categories like quads. Instead, quads are highly variable  and come from a wide range of brands, so it is harder for the model to learn which set of features designate a quadrilateral. Since a large number of features are then associated with quads, the model gets more easily confused. When it sees some of those same features in other classes, it tries to guess quad. This is further explored in the next section.}

\mt{That was a really quick and dirty version of what I'm trying to say... Am I missing any big points, before I try to expand it and clean up the technical part of the writing?}

@@

Heatmaps
  \begin{itemize}
  \item Verify that the model is using appropriate features when predictions are correct
  \item Show examples for confident misclassifications (like Seychelles), but make the case that the mistakes are understandable
  \end{itemize}
